Inscrutable mysteries underlying the horizon of our journey. Are often shady the paths that we must go. Many of our experiences stained our memories of greyish colors with ink that can not be changed. Still, certain tones, certain sounds, remain rooted in the foundations of memory working as anchors that prevent completely and eternally sink us. Some faces, some voices, some smiles, some advice, or an occasional correction are always pleasant company in our memories and encouragement when it comes to achieving the goals, to undertake new challenges and face major changes. 

Thank God, eternal force that has design and order to this universe, so wonderful and mysterious, which again and again provides us with concerns and mysteries able to capture, inundate and complement the restless mind with which we have been provided.

Gracias a Gloria Calabrán Contreras, amante, esposa y conviviente civil. Por compartir nuestro dia a dia. Por nuestras sobrinas. Por la paciencia y apoyo. Por la vida en conjunto, en la que dia a dia trabajamos como equipo.
%Thanks to my exfiancee, Carolina, for their endless support and patience, having been close to me in this way. By converting their desires and my desires in our dreams and goals. By having their life and purposes jointly with mine, forever hear my stories nonsense on relativity, quantum, philosophy and much more, always paying a tender and sincere interest. 
Thanks grandma and aunt for host and support when the rest of the universe looked in another direction, thanks a lot.

\small{
I strongly thank my advisors, Dr Jorge Silva and Dr Julian Ortiz, and to Dr Milan Derpich, who have allowed me to develop and formulate innovative and relevant concepts. Thanks for a collaborative thesis in areas of great importance to me, and for your invaluable support and for the various tasks entrusted. Thanks Thomas Peet and Ivan Castro and Jorge Toledo by helping and sharing with me along the formalization of subjects for this work.
}

Special thanks to The Coquetos and its support without which this dissertation would not have been possible. Muchas gracias Srta Andrea, por sus animadas canciones sin fin. Gracias por su chaleco verde que mantenemos en IDS y que nos acompaña a nuestras salidas, y que por sobre todo nos inspira a mejorar cada dia y a rectificar continuamente nuestros infinitos errores. Un abrazo al Gran Sebita PRESO (\&GOD), ayudante incansable de las futuras generaciones, amante de la enseñaza y la educacion temprana. Alma noble preocupada por estudiantas, mechonas, y exmechonas de tantas generaciones. Como olvidar sus regalos misteriosos y las hordas de estudiantas esperando sus sabios consejos y su coqueta sonrisa y mirada. I hope one day to be able to see you two again, to embrace you both, to kiss both of you, and to hear her song "Abuelito dime tú".

\small{
Finally, to you, that for one reason or another uses this thesis,  ¡Hasta Siempre!. 
\begin{flushright}
Felipe Andr\'es Santib\'añez Leal
\end{flushright}
}